---
title: "Addressing Gender and Disability Gaps at ISED"
author: "Hyoeun Park, Jaffa Romain"
date: "Sunday, February 6, 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(readr)
library(tidyverse)
ee <- read.csv("subset-3-sous-ensemble-3.csv")
```

# Addressing Gender and Disability Gaps at ISED

## Gap between female/gender diverse and male individuals

```{r gender_graph}
ee <- ee %>% mutate(Gender = case_when(
  DESCRIP_E == "Male gender" ~ "Male",
  DESCRIP_E == "Female gender" | DESCRIP_E == "Gender diverse" ~ 
    "Female/Gender diverse"
))

gender <- ee %>%  
  select(Gender, QUESTION, BYCOND, SCORE100, DEPT_E, ANSWER1,
         ANSWER2, ANSWER3, ANSWER4, ANSWER5, ANSWER6, ANSWER7) %>% 
  filter(SCORE100 != 9999) %>% 
  filter(DEPT_E == "Innovation, Science and Economic Development Canada") %>% 
  na.exclude()

gender_diff <- gender %>% 
  group_by(QUESTION, Gender) %>% 
  summarize(mean = mean(SCORE100, na.rm = TRUE)) %>%  
  pivot_wider(names_from = Gender,  values_from = mean) %>% 
  summarize(diff = `Female/Gender diverse` - `Male`)

gender_diff %>% 
  ggplot(aes(x = diff, colour = diff)) + 
  geom_histogram(fill = "orange") + 
  theme_minimal() +
  labs(x = "Difference",
       y = "Count",
       caption = "Figure 1: Difference between Females/Gender Diverse and Males \n (The higher the difference in absolute value, the higher the disparity)")
```

Figure 1 captures how the scores of each questions differ for female/gender diverse individuals to male individuals. For each of the differences, we subtracted the score of female/gender diverse individuals by scores of males. A higher difference in absolute values indicate that the disparity is much larger for the scores between the two subgroups. In Figure 1, we observe that most values lie below 0. Furthermore, most differences are centered around -5.   

```{r gender_test}
t.test(gender_diff$diff, 
       var.equal = T, alternative = "less")
```

Since we have noticed some plausible differences in Figure 1, we can conduct a t-test to figure out whether the difference of scores between the two groups are equal to 0 or less than 0. As a result, we have strong evidence against the hypothesis that the scores between males and females are equal to 0. Hence, we conclude that the difference in scores are less than zero for males and females/gender diverse employees.

Now, let us investigate some of the survey questions to see what might be the underlying causes of such difference in scores.

```{r gender_promo_retention}
# Question 42. I believe I have opportunities 
# for promotion within my department or agency, 
# given my education, skills and experience. 
gender_promo <- gender %>% filter(QUESTION == "Q42")
Q42_total <- cbind(gender_promo %>%  
                     group_by(Gender) %>% 
                     summarise(Count = c(sum(ANSWER1), sum(ANSWER2),
                                         sum(ANSWER3), sum(ANSWER4),
                                         sum(ANSWER5), sum(ANSWER6))),
                 Answer = 
                   rep(c("Strongly Agree","Somewhat agree",
                                    "Neither agree not disagree",
                                    "Somewhat disagree",
                                    "Strongly disagree",
                                    "Don't know"), 2))
gender_42 <- Q42_total %>% ggplot(aes(x = Gender, fill = Answer, y = Count)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  labs(caption = "Figure 1.1: Opinions of promotion by female/gender diverse and male groups") + 
  labs(x = "Gender", y = "Count")


# Question 53. Do you intend to leave 
# your current position in the next two years?
gender_leave <- ee %>% 
  filter(QUESTION == "Q53" & 
           DEPT_E == "Innovation, Science and Economic Development Canada") %>%
  select(Gender, AGREE, ANSWER1, ANSWER2, ANSWER3) %>% 
  na.exclude()

Q53_total <- rbind(gender_leave %>% 
      group_by(Gender) %>% 
      summarize(Count = sum(AGREE)),
      gender_leave %>% 
      group_by(Gender) %>% 
      summarize(Count = sum(ANSWER2)),
      gender_leave %>% 
      group_by(Gender) %>% 
      summarize(Count = sum(ANSWER3)))
Q53_total <- Q53_total %>% mutate(Answer = c(rep("Yes", 2), rep("No", 2),
                                             rep("Unsure", 2)))

gender_53 <- Q53_total %>% ggplot(aes(x = Gender, fill = Answer, y = Count)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal() +
  labs(caption = "Figure 1.2: Number of people who expressed their potential to leave within 2 years.")

gender_42
gender_53
```

Based on Questions 42 and 53, we were able to visualize the distribution of answers across the two groups as Figures 1.1 and 1.2. First of all, Figure 1.1 shows us that the answers were similar across the two groups. In fact, a lot of employees in both groups strongly agreed and strongly disagreed at the same time regarding their promotion based on skills, experience, and education.

Meanwhile, we see in Figure 1.2 that several people in the female/gender diverse group strongly expressed their intent to leave ISED within the next 2 years in comparison to the proportion in male individuals. Also, a lot of female/gender diverse individuals were unsure of leaving within the 2 years, which could imply that we need to further investigate what might be the cause of it. 

```{r gender_harass}
# Question 55: Having carefully read the definition of harassment, 
# have you been the victim of harassment on the job in the past 12 months? 
gender_harass <- ee %>% 
  filter(QUESTION == "Q55" & 
           DEPT_E == "Innovation, Science and Economic Development Canada") %>%
  select(Gender, AGREE, ANSWER2) %>% 
  na.exclude()
Q55_total <- rbind(gender_harass %>% 
                     group_by(Gender) %>% 
                     summarise(Count = sum(AGREE)),
                   gender_harass %>% 
                     group_by(Gender) %>% 
                     summarise(Count = sum(ANSWER2)))
Q55_total <- Q55_total %>% mutate(Answer = c(rep("Yes", 2), rep("No", 2)))

# Number of people who've experienced harassment(s) over the past one year
gender_55 <- Q55_total %>% ggplot(aes(x = Gender, fill = Answer, y = Count)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal() +
  labs(caption = "Figure 1.3: Number of people who have experienced harassment in the workplace.")

# Question 57: Type of harassment
type_harass <- ee %>% 
  filter(QUESTION == "Q57a" | QUESTION == "Q57b" |
           QUESTION == "Q57c" | QUESTION == "Q57d" |
           QUESTION == "Q57e" | QUESTION == "Q57f" |
           QUESTION == "Q57g" | QUESTION == "Q57h" |
           QUESTION == "Q57i" | QUESTION == "Q57j" |
           QUESTION == "Q57k" | QUESTION == "Q57l" |
           QUESTION == "Q57m") %>% 
  filter(DEPT_E == "Innovation, Science and Economic Development Canada") %>% 
  select(Gender, AGREE, QUESTION) %>% 
  na.exclude() %>% 
  group_by(Gender, QUESTION) %>% 
  summarize(Total = sum(AGREE)) %>% 
  mutate(Type = case_when(
    QUESTION == "Q57a" ~ "Aggressive behaviour",
    QUESTION == "Q57b" ~ "Excessive control",
    QUESTION == "Q57c" ~ "Being excluded or being ignored",
    QUESTION == "Q57d" ~ "Humiliation",
    QUESTION == "Q57e" ~ "Interference with work or withholding resources",
    QUESTION == "Q57f" ~ "Offensive remark",
    QUESTION == "Q57g" ~ "Personal attack",
    QUESTION == "Q57i" ~ "Sexual comment or gesture",
    QUESTION == "Q57j" ~ "Threat",
    QUESTION == "Q57k" ~ "Unfair treatment",
    QUESTION == "Q57l" ~ "Yelling or shouting",
    QUESTION == "Q57m" ~ "Other"
  )) %>% 
  select(-QUESTION)

# Plot: Counts of each type of harassment
gender_57 <- type_harass %>% 
  ggplot(aes(x = Gender, y = Total, fill = Type)) + 
  geom_bar(stat = 'identity', position = position_dodge()) +
  theme_minimal() + 
  labs(caption = "Figure 1.4: Counting the number of harassments experienced by gender.")

# Question 60: I am satisfied with how matters 
# related to harassment are resolved in my department or agency. 
harass_satisfy <- ee %>% 
  filter(QUESTION == "Q60" &
           DEPT_E == "Innovation, Science and Economic Development Canada") %>%
  select(Gender, ANSWER1, ANSWER2, ANSWER3, ANSWER4, ANSWER5, ANSWER6) %>% 
  na.exclude()
Q60_total <- cbind(harass_satisfy %>% 
                     group_by(Gender) %>% 
                     summarise(Count = c(sum(ANSWER1), 
                                         sum(ANSWER2),
                                         sum(ANSWER3),
                                         sum(ANSWER4),
                                         sum(ANSWER5),
                                         sum(ANSWER6))),
                   Answer = 
                     rep(c("Strongly Agree","Somewhat agree",
                                    "Neither agree not disagree",
                                    "Somewhat disagree",
                                    "Strongly disagree",
                                    "Don't know"), 2))

# Plot: How people are satisfied with harassment issues being resolved
gender_60 <- Q60_total %>% ggplot(aes(x = Gender, y = Count, fill = Answer)) + 
  geom_bar(stat = 'identity', position = position_dodge()) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2") +
  labs(caption = "Figure 1.5: Responses to how people are satisfied with harassment issues being resolved, based on gender.")

gender_55
gender_57
gender_60
```

According to Figure 1.3 which was based on Question 55 regarding the experience of harassment, it seems like both groups have experienced a similar proportion of harassment in the workplace. 

When looking at the types of harassment they have experienced (Figure 1.4), we see that the distribution is also quite similar across the two groups. However, We notice that female/gender diverse individuals experienced more excessive control, interference with work or withholding resources and sexual comment or gestures. On the other hand, male workers faced more personal attack and unfair treatment. 

Finally, Figure 1.5 shows how people are satisfied with harassment issues being resolved. In general, many people in the female/gender diverse category were neutral. Moreover, there were more female/gender diverse individuals who somewhat disagreed with the resolutions compared to male workers. A lot of male individuals responded "Don't know" followed by "Strongly agree" as the second most frequent response. Regarding the fact that many responses were neutral, we can infer that people are not fully satisfied with the harassment issues being resolved.

Based on such results, we could possibly give a suggestion to 


## Gap between people with disabilities and no disabilities


```{r}
ee <- ee %>% mutate(disability_health = case_when(
  ee$DESCRIP_E == "Person with a disability" | 
  ee$DESCRIP_E == "A seeing disability affects vision, including total blindness, partial sight and visual distortion" |
    ee$DESCRIP_E == "A hearing disability affects ability to hear, including being hard of hearing, deafness or acoustic distortion" |
    ee$DESCRIP_E == "A mobility issue affects ability to move your body, including the required use of a wheelchair or a cane, or other issues impacting your mobility" | 
    ee$DESCRIP_E == "An issue with flexibility or dexterity affects ability to move joints or perform motor tasks, especially with your hands" |
    ee$DESCRIP_E == "A mental health issue affects psychology or behaviour, such as anxiety, depression or social / compulsive disorder or phobia or psychiatric illness" |
    ee$DESCRIP_E == "A sensory / environmental disability affects sensitivity to light, sounds or other distractions, as well as allergens and other environmental sensitivities" |
    ee$DESCRIP_E == "A chronic health condition or pain affects ability to function on a regular or episodic basis due to migraines, Crohn's disease, colitis, and other disabilities or health conditions" |
    ee$DESCRIP_E == "A cognitive disability affects ability to carry out tasks involving executive functioning, such as planning and organization, learning information, communication and memory, including autism or Asperger's syndrome, attention deficit disorder, and learning disabilities" |
    ee$DESCRIP_E == "An intellectual disability affects your ability to learn and to adapt behaviour to different situations" |
    ee$DESCRIP_E == "Other disability (including: learning disabilities, developmental disabilities and all other types of disabilities)" ~ "Disability"
    ,
  str_detect(ee$DESCRIP_E, "Not a person with a disability") ~ 
    "No disabilities"
))

disability_health <- ee %>% 
  select(disability_health, QUESTION, BYCOND, SCORE100, DEPT_E,
         ANSWER1, ANSWER2, ANSWER3, ANSWER4, ANSWER5, ANSWER6, ANSWER7) %>% 
  filter(SCORE100 != 9999) %>%
  filter(DEPT_E == "Innovation, Science and Economic Development Canada")

dis_diff <- disability_health %>% 
  group_by(QUESTION, disability_health) %>% 
  summarize(mean = mean(SCORE100, na.rm = TRUE)) %>%  
  pivot_wider(names_from = disability_health,  values_from = mean) %>% 
  summarize(diff = `Disability` - `No disabilities`)

dis_diff %>% 
  ggplot(aes(x = diff, colour = diff)) + 
  geom_histogram(fill = "green") + 
  theme_minimal() +
  labs(x = "Difference",
       y = "Count",
       caption = "Figure 3: Difference between people with disabilities/health concerns and those with no disabilities \n (A smaller difference in absolute value indicate a smaller gap)")
```

```{r disability_test}
t.test(dis_diff$diff, alternative = "less", var.equal = T)
```

We have strong evidence that the score between males and females are different than 0.

```{r disability_promo_leave}
# Question 42. I believe I have opportunities 
# for promotion within my department or agency, 
# given my education, skills and experience. 
dis_promo <- disability_health %>% filter(QUESTION == "Q42") %>% na.exclude()
Q42_dis <- cbind(dis_promo %>%  
                     group_by(disability_health) %>% 
                     summarise(Count = c(sum(ANSWER1), sum(ANSWER2),
                                         sum(ANSWER3), sum(ANSWER4),
                                         sum(ANSWER5), sum(ANSWER6))),
                 Answer = 
                   rep(c("Strongly Agree","Somewhat agree",
                                    "Neither agree not disagree",
                                    "Somewhat disagree",
                                    "Strongly disagree",
                                    "Don't know"), 2))
dis_42 <- Q42_dis %>% ggplot(aes(x = disability_health, fill = Answer, y = Count)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  labs(caption = "Figure 2.1: Opinions of promotion by disability and non-disability groups") + 
  labs(x = "Disability", y = "Count")


# Question 53. Do you intend to leave 
# your current position in the next two years?
dis_leave <- ee %>% 
  filter(QUESTION == "Q53" & 
           DEPT_E == "Innovation, Science and Economic Development Canada") %>%
  select(disability_health, AGREE, ANSWER1, ANSWER2, ANSWER3) %>% 
  na.exclude()

Q53_dis <- cbind(dis_leave %>% 
      group_by(disability_health) %>% 
      summarize(Count = c(sum(AGREE), sum(ANSWER2), sum(ANSWER3))),
      Answer = rep(c("Yes", "No", "Unsure"), 2))

dis_53 <- Q53_dis %>% ggplot(aes(x = disability_health, fill = Answer, y = Count)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal() +
  labs(caption = "Figure 2.2: Number of people who expressed their potential to leave within the next 2 years.", 
       x = "Disability")

dis_42
dis_53
```


```{r dis_questions}
dis_aftercovid <- disability_health %>% 
  filter(QUESTION == "Q91") %>% 
  na.exclude
Q91_dis <- cbind(dis_aftercovid %>%  
                     group_by(disability_health) %>% 
                     summarise(Count = c(sum(ANSWER1), sum(ANSWER2),
                                         sum(ANSWER3), sum(ANSWER4),
                                         sum(ANSWER5), sum(ANSWER6))),
                 Answer = 
                   rep(c("Strongly Agree","Somewhat agree",
                                    "Neither agree not disagree",
                                    "Somewhat disagree",
                                    "Strongly disagree",
                                    "Don't know"), 2))
dis_91 <- Q91_dis %>% ggplot(aes(x = disability_health, fill = Answer, y = Count)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  labs(caption = "Figure 2.3: Whether people are satisfied with the workplace accommodation measures that were implemented.") + 
  labs(x = "Disability", y = "Count")
dis_91

dis_Q92 <- disability_health %>% 
  filter(QUESTION == "Q92") %>% 
  na.exclude
Q92_dis <- cbind(dis_Q92 %>%  
                     group_by(disability_health) %>% 
                     summarise(Count = c(sum(ANSWER1), sum(ANSWER2),
                                         sum(ANSWER3), sum(ANSWER4),
                                         sum(ANSWER5), sum(ANSWER6))),
                 Answer = 
                   rep(c("Strongly Agree","Somewhat agree",
                                    "Neither agree not disagree",
                                    "Somewhat disagree",
                                    "Strongly disagree",
                                    "Don't know"), 2))
dis_92 <- Q92_dis %>% ggplot(aes(x = disability_health, fill = Answer, y = Count)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  labs(caption = "Figure 2.4: How people are comfortable with requesting workplace accommodation measures from their immediate supervisor.") + 
  labs(x = "Disability", y = "Count")
dis_92

Q70q_dis <- disability_health %>% filter(QUESTION == "Q70q") %>% 
  na.exclude() %>% 
  group_by(disability_health) %>% 
  summarise(Count = c(sum(ANSWER1), sum(ANSWER2), sum(ANSWER3), 
                      sum(ANSWER4), sum(ANSWER5), sum(ANSWER6))) %>% 
  cbind(Answer = rep(c("Strongly Agree","Somewhat agree",
                                    "Neither agree not disagree",
                                    "Somewhat disagree",
                                    "Strongly disagree",
                                    "Don't know"), 2))
dis_70q <- Q70q_dis %>% ggplot(aes(x = disability_health, y = Count, fill = Answer)) +
  geom_bar(position = position_dodge(), stat = 'identity') +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  labs(caption = "Figure 2.5: Number of people who feel stressed in the workplace due to accessibility or accommodation issues")

dis_70q
```
