---
title: "Addressing Gender Gap at ISED"
author: "Hyoeun Park"
date: "February 6, 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(readr)
library(tidyverse)
ee <- read.csv("subset-3-sous-ensemble-3.csv")
```

# Addressing Gender Gap at ISED

## Gap between female/gender diverse and male individuals

```{r gender_graph}
ee <- ee %>% mutate(Gender = case_when(
  DESCRIP_E == "Male gender" ~ "Male",
  DESCRIP_E == "Female gender" | DESCRIP_E == "Gender diverse" ~ 
    "Female/Gender diverse"
))

gender <- ee %>%  
  select(Gender, QUESTION, BYCOND, SCORE100, DEPT_E, ANSWER1,
         ANSWER2, ANSWER3, ANSWER4, ANSWER5, ANSWER6, ANSWER7) %>% 
  filter(SCORE100 != 9999) %>% 
  filter(DEPT_E == "Innovation, Science and Economic Development Canada") %>% 
  na.exclude()

gender_diff <- gender %>% 
  group_by(QUESTION, Gender) %>% 
  summarize(mean = mean(SCORE100, na.rm = TRUE)) %>%  
  pivot_wider(names_from = Gender,  values_from = mean) %>% 
  summarize(diff = `Female/Gender diverse` - `Male`)

gender_diff %>% 
  ggplot(aes(x = diff, colour = diff)) + 
  geom_histogram(fill = "orange") + 
  theme_minimal() +
  labs(x = "Difference",
       y = "Count",
       caption = "Figure 1: Difference between Females/Gender Diverse and Males \n (The higher the difference in absolute value, the higher the disparity)")
```

Figure 1 captures how the scores of each questions differ for female/gender diverse individuals to male individuals. For each of the differences, we subtracted the score of female/gender diverse individuals by scores of males. A higher difference in absolute values indicate that the disparity is much larger for the scores between the two subgroups. In Figure 1, we observe that most values lie below 0. Furthermore, most differences are centered around -5.   

```{r gender_test}
t.test(gender_diff$diff, 
       var.equal = T, alternative = "less")
```

Since we have noticed some plausible differences in Figure 1, we can conduct a t-test to figure out whether the difference of scores between the two groups are equal to 0 or less than 0. As a result, we have strong evidence against the hypothesis that the scores between males and females are equal to 0. Hence, we conclude that the difference in scores are less than zero for males and females/gender diverse employees.

Now, let us investigate some of the survey questions to see what might be the underlying causes of such difference in scores.

```{r gender_promo_retention}
# Question 42. I believe I have opportunities 
# for promotion within my department or agency, 
# given my education, skills and experience. 
gender_promo <- gender %>% filter(QUESTION == "Q42")
Q42_total <- cbind(gender_promo %>%  
                     group_by(Gender) %>% 
                     summarise(Count = c(sum(ANSWER1), sum(ANSWER2),
                                         sum(ANSWER3), sum(ANSWER4),
                                         sum(ANSWER5), sum(ANSWER6))),
                 Answer = 
                   rep(c("Strongly Agree","Somewhat agree",
                                    "Neither agree not disagree",
                                    "Somewhat disagree",
                                    "Strongly disagree",
                                    "Don't know"), 2))
gender_42 <- Q42_total %>% ggplot(aes(x = Gender, fill = Answer, y = Count)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  labs(caption = "Figure 1.1: Opinions of promotion by female/gender diverse and male groups") + 
  labs(x = "Gender", y = "Count")


# Question 53. Do you intend to leave 
# your current position in the next two years?
gender_leave <- ee %>% 
  filter(QUESTION == "Q53" & 
           DEPT_E == "Innovation, Science and Economic Development Canada") %>%
  select(Gender, AGREE, ANSWER1, ANSWER2, ANSWER3) %>% 
  na.exclude()

Q53_total <- rbind(gender_leave %>% 
      group_by(Gender) %>% 
      summarize(Count = sum(AGREE)),
      gender_leave %>% 
      group_by(Gender) %>% 
      summarize(Count = sum(ANSWER2)),
      gender_leave %>% 
      group_by(Gender) %>% 
      summarize(Count = sum(ANSWER3)))
Q53_total <- Q53_total %>% mutate(Answer = c(rep("Yes", 2), rep("No", 2),
                                             rep("Unsure", 2)))

gender_53 <- Q53_total %>% ggplot(aes(x = Gender, fill = Answer, y = Count)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal() +
  labs(caption = "Figure 1.2: Number of people who expressed their potential to leave within 2 years.")

gender_42
gender_53
```

Based on Questions 42 and 53, we were able to visualize the distribution of answers across the two groups as Figures 1.1 and 1.2. First of all, Figure 1.1 shows us that the answers were similar across the two groups. In fact, a lot of employees in both groups strongly agreed and strongly disagreed at the same time regarding their promotion based on skills, experience, and education.

Meanwhile, we see in Figure 1.2 that several people in the female/gender diverse group strongly expressed their intent to leave ISED within the next 2 years in comparison to the proportion in male individuals. Also, a lot of female/gender diverse individuals were unsure of leaving within the 2 years, which could imply that we need to further investigate what might be the cause of it. 

```{r gender_harass}
# Question 55: Having carefully read the definition of harassment, 
# have you been the victim of harassment on the job in the past 12 months? 
gender_harass <- ee %>% 
  filter(QUESTION == "Q55" & 
           DEPT_E == "Innovation, Science and Economic Development Canada") %>%
  select(Gender, AGREE, ANSWER2) %>% 
  na.exclude()
Q55_total <- rbind(gender_harass %>% 
                     group_by(Gender) %>% 
                     summarise(Count = sum(AGREE)),
                   gender_harass %>% 
                     group_by(Gender) %>% 
                     summarise(Count = sum(ANSWER2)))
Q55_total <- Q55_total %>% mutate(Answer = c(rep("Yes", 2), rep("No", 2)))

# Number of people who've experienced harassment(s) over the past one year
gender_55 <- Q55_total %>% ggplot(aes(x = Gender, fill = Answer, y = Count)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal() +
  labs(caption = "Figure 1.3: Number of people who have experienced harassment in the workplace.")

# Question 57: Type of harassment
type_harass <- ee %>% 
  filter(QUESTION == "Q57a" | QUESTION == "Q57b" |
           QUESTION == "Q57c" | QUESTION == "Q57d" |
           QUESTION == "Q57e" | QUESTION == "Q57f" |
           QUESTION == "Q57g" | QUESTION == "Q57h" |
           QUESTION == "Q57i" | QUESTION == "Q57j" |
           QUESTION == "Q57k" | QUESTION == "Q57l" |
           QUESTION == "Q57m") %>% 
  filter(DEPT_E == "Innovation, Science and Economic Development Canada") %>% 
  select(Gender, AGREE, QUESTION) %>% 
  na.exclude() %>% 
  group_by(Gender, QUESTION) %>% 
  summarize(Total = sum(AGREE)) %>% 
  mutate(Type = case_when(
    QUESTION == "Q57a" ~ "Aggressive behaviour",
    QUESTION == "Q57b" ~ "Excessive control",
    QUESTION == "Q57c" ~ "Being excluded or being ignored",
    QUESTION == "Q57d" ~ "Humiliation",
    QUESTION == "Q57e" ~ "Interference with work or withholding resources",
    QUESTION == "Q57f" ~ "Offensive remark",
    QUESTION == "Q57g" ~ "Personal attack",
    QUESTION == "Q57i" ~ "Sexual comment or gesture",
    QUESTION == "Q57j" ~ "Threat",
    QUESTION == "Q57k" ~ "Unfair treatment",
    QUESTION == "Q57l" ~ "Yelling or shouting",
    QUESTION == "Q57m" ~ "Other"
  )) %>% 
  select(-QUESTION)

# Plot: Counts of each type of harassment
gender_57 <- type_harass %>% 
  ggplot(aes(x = Gender, y = Total, fill = Type)) + 
  geom_bar(stat = 'identity', position = position_dodge()) +
  theme_minimal() + 
  labs(caption = "Figure 1.4: Counting the number of harassments experienced by gender.")

# Question 60: I am satisfied with how matters 
# related to harassment are resolved in my department or agency. 
harass_satisfy <- ee %>% 
  filter(QUESTION == "Q60" &
           DEPT_E == "Innovation, Science and Economic Development Canada") %>%
  select(Gender, ANSWER1, ANSWER2, ANSWER3, ANSWER4, ANSWER5, ANSWER6) %>% 
  na.exclude()
Q60_total <- cbind(harass_satisfy %>% 
                     group_by(Gender) %>% 
                     summarise(Count = c(sum(ANSWER1), 
                                         sum(ANSWER2),
                                         sum(ANSWER3),
                                         sum(ANSWER4),
                                         sum(ANSWER5),
                                         sum(ANSWER6))),
                   Answer = 
                     rep(c("Strongly Agree","Somewhat agree",
                                    "Neither agree not disagree",
                                    "Somewhat disagree",
                                    "Strongly disagree",
                                    "Don't know"), 2))

# Plot: How people are satisfied with harassment issues being resolved
gender_60 <- Q60_total %>% ggplot(aes(x = Gender, y = Count, fill = Answer)) + 
  geom_bar(stat = 'identity', position = position_dodge()) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2") +
  labs(caption = "Figure 1.5: Responses to how people are satisfied with harassment issues being resolved, based on gender.")

gender_55
gender_57
gender_60
```

According to Figure 1.3 which was based on Question 55 regarding the experience of harassment, it seems like both groups have experienced a similar proportion of harassment in the workplace. 

When looking at the types of harassment they have experienced (Figure 1.4), we see that the distribution is also quite similar across the two groups. However, We notice that female/gender diverse individuals experienced more excessive control, interference with work or withholding resources and sexual comment or gestures. On the other hand, male workers faced more personal attack and unfair treatment. 

Finally, Figure 1.5 shows how people are satisfied with harassment issues being resolved. In general, many people in the female/gender diverse category were neutral. Moreover, there were more female/gender diverse individuals who somewhat disagreed with the resolutions compared to male workers. A lot of male individuals responded "Don't know" followed by "Strongly agree" as the second most frequent response. Regarding the fact that many responses were neutral, we can infer that people are not fully satisfied with the harassment issues being resolved.

### Conclusion

Based on the survey responses and the dataset that we were given, it is difficult to determine the number of female/gender diverse employees and male employees at ISED. We can only tell that there were more female/gender diverse individuals who responded to the survey. If a lot of female/gender diverse individuals decided to answer questions more than males in order to raise their voices on inequity in the workplace, this could mean that we might need to work more on promoting gender equity / equality in the workplace.

We could possibly give a few suggestions to ISED using the results above. First, they can foster gender equity / equality starting with the recruitment process. Hiring more EE (Employment equity) groups will lead to more equal outcomes. During the recruitment process, ISED should make sure recruitment managers go through adequate training sessions to reduce internal bias towards job candidates regarding their gender. They can also diversify hiring managers. For instance, they can have an equal/similar number of men, women, gender diverse hiring managers. Finally, when it comes to harassment issues, the company should hold people accountable such as the offender of harassment and the manager. It is essential that all individuals in the workplace go through adequate hours of training to reduce harassment issues.
