---
title: "CANDEV- ISED ANALYSIS"
author: "Jaffa Romain"
date: "05/02/2022"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
library(kableExtra)
```

# Data Cleaning
```{r data}
CANDEV_data <- read_csv("data/subset-3-sous-ensemble-3.csv")
CANDEV_data <- CANDEV_data %>% filter(SCORE100!= 9999)

ISED_only <- CANDEV_data %>% filter(DEPT_E == "Innovation, Science and Economic Development Canada")

# Visible minority
VM_vs_NVM <- ISED_only %>% filter(BYCOND == "Q120 = 1" | BYCOND == "Q120 = 2")
VM_vs_NVM %>%  group_by(DESCRIP_E) %>% count()

Ethnic_Groups <- ISED_only %>% filter(BYCOND == "Q121A = 1" |BYCOND ==  "Q121B = 1"|BYCOND == "Q121C = 1" |BYCOND == "Q121D = 1"|BYCOND =="Q121E = 1"|BYCOND == "Q121F = 1"|BYCOND == "Q121G = 1"|BYCOND == "Q121H = 1"|BYCOND ==  "Q121I = 1"| BYCOND == "Q121J = 1"| BYCOND =="Q121K = 1" |BYCOND =="Q120 = 2")
Ethnic_Groups %>%  group_by(DESCRIP_E) %>% count()

# Indigenous
IvsNI <- ISED_only %>% filter(BYCOND == "Q116 = 1" | BYCOND ==  "Q116 = 2")

# Gender
gender <- ISED_only %>% filter(BYCOND == "Q115 = 1" | BYCOND ==  "Q115 = 2"|BYCOND == "Q115 = 3")


# Disabilities

access <- ISED_only %>% filter(BYCOND == "Q118 = 1" | BYCOND ==  "Q118 = 2")

```

# ANOVA assessing difference in groups
 - score100 is used as measure

## Minority vs Non-minority
```{r}
# RACE - questions related to cultural diversity and racism
minorities <- VM_vs_NVM %>% filter(QUESTION == "Q22"| QUESTION == "Q45"| QUESTION == "Q46"| QUESTION == "Q47"| QUESTION == "Q48"| QUESTION == "6Q4a" | QUESTION == "Q64c" | QUESTION == "6Q4b"| QUESTION == "Q69") 

t.test(SCORE100~DESCRIP_E, data = minorities)


# which questions/prompts have the biggest difference in scores?


differencesminorities<- minorities %>% group_by(QUESTION, BYCOND) %>% summarize(mean = mean(SCORE100, na.rm = TRUE)) %>%  pivot_wider(names_from = BYCOND,  values_from = mean) %>% summarize(diff = `Q120 = 1` - `Q120 = 2`) 

differencesminorities


minorities %>%  ggplot(aes(x = DESCRIP_E, y = SCORE100)) + geom_boxplot(
  aes(fill = DESCRIP_E),
  position = position_dodge(0.9) 
  ) +
  scale_fill_manual(values = c("#999999", "#E69F00"))

```

- # significant difference between cultural scores for minority vs minority (p < 0.05)
_ what are the differences?
- minority groups are rated lower for all of the questions
- the biggest gap in scores among these questions is q48: Question 48. I would feel comfortable sharing concerns about issues related to racism in the workplace with a person of authority (e.g., immediate supervisor, senior manager, ombudsman, human resources advisor).

- second highest: Q25
- difference in average score (out of 100) is 12.

How comfortable do employees feel sharing concerns?
```{r}
minorities %>% group_by(QUESTION, BYCOND) %>% summarize(mean = mean(SCORE100, na.rm = TRUE))


```


# Indigenous vs. non-Indigenous
```{r}
IN <- IvsNI %>% filter(QUESTION == "Q22"| QUESTION == "Q45"| QUESTION == "Q46"| QUESTION == "Q47"| QUESTION == "Q48"| QUESTION == "6Q4a" | QUESTION == "Q64c" | QUESTION == "6Q4b"| QUESTION == "Q69") 

t.test(SCORE100~DESCRIP_E, data =IN)


# which questions/prompts have the biggest difference in scores?


differencesI<-IN %>% group_by(QUESTION, BYCOND) %>% summarize(mean = mean(SCORE100, na.rm = TRUE)) %>%  pivot_wider(names_from = BYCOND,  values_from = mean) %>% summarize(diff = `Q116 = 1` - `Q116 = 2`) 




IN %>%  ggplot(aes(x = DESCRIP_E, y = SCORE100)) + geom_boxplot(
  aes(fill = DESCRIP_E),
  position = position_dodge(0.9) 
  ) +
  scale_fill_manual(values = c("#999999", "#E69F00"))
```
- SAME AS MINORITY VS NON-MINORITY - BUT DIFFERENCE NOT STATISTICALLY SIGNIFICANT



## all groups - comparing scores of groups based on organization threshold
```{r}
Ethnic_Groups <- CANDEV_data %>% filter(BYCOND == "Q121A = 1" |BYCOND ==  "Q121B = 1"|BYCOND == "Q121C = 1" |BYCOND == "Q121D = 1"|BYCOND =="Q121E = 1"|BYCOND == "Q121F = 1"|BYCOND == "Q121G = 1"|BYCOND == "Q121H = 1"|BYCOND ==  "Q121I = 1"| BYCOND == "Q121J = 1"| BYCOND =="Q121K = 1" |BYCOND =="Q120 = 2") %>% filter(SCORE100!= 9999) %>% filter(QUESTION == "Q22"| QUESTION == "Q45"| QUESTION == "Q46"| QUESTION == "Q47"| QUESTION == "Q48"| QUESTION == "6Q4a" | QUESTION == "Q64c" | QUESTION == "6Q4b"| QUESTION == "Q69") 

total_average = mean(Ethnic_Groups$SCORE100)


ALLGroups <- ISED_only %>% filter(BYCOND == "Q121A = 1" |BYCOND ==  "Q121B = 1"|BYCOND == "Q121C = 1" |BYCOND == "Q121D = 1"|BYCOND =="Q121E = 1"|BYCOND == "Q121F = 1"|BYCOND == "Q121G = 1"|BYCOND == "Q121H = 1"|BYCOND ==  "Q121I = 1"| BYCOND == "Q121J = 1"| BYCOND =="Q121K = 1" |BYCOND =="Q120 = 2") %>% filter(SCORE100!= 9999) %>% filter(QUESTION == "Q22"| QUESTION == "Q45"| QUESTION == "Q46"| QUESTION == "Q47"| QUESTION == "Q48"| QUESTION == "6Q4a" | QUESTION == "Q64c" | QUESTION == "6Q4b"| QUESTION == "Q69") 

ALLGroups <- ALLGroups %>% mutate(Ethnicity = case_when(DESCRIP_E == "Black"~"Black",  
                                                       DESCRIP_E == "Chinese"~"Asian",
                                                       DESCRIP_E == "Filipino"~"Asian",
                                                       DESCRIP_E == "Non-visible minority"~"Non-visible minority",
                                                       DESCRIP_E == "Non-White Latin American (including: Indigenous persons from Central and South America, etc.)"~"Latin American",
                                                       DESCRIP_E == "Non-White West Asian, North African or Arab (including: Egyptian; Libyan; Lebanese; Iranian; etc.)"~"West Asian/North African/Arab",
                                                       DESCRIP_E == "Other visible minority group"~"Other",
                                                       DESCRIP_E == "Person of mixed origin (with one parent in one of the visible minority groups)"~"Mixed",
                                                       DESCRIP_E == "South Asian/East Indian (including: Indian from India; Bangladeshi; Pakistani; East Indian from Guyana, Trinidad, East Africa; etc.)"~"South Asian/East Indian",
                                                       DESCRIP_E == "Southeast Asian (including: Burmese; Cambodian; Laotian; Thai; Vietnamese; etc.)"~"Southeast Asian"))

ALLGroups %>% group_by(Ethnicity) %>% summarize(mean = mean(SCORE100, na.rm = TRUE)) %>% ggplot(aes(x = Ethnicity, y = mean)) + geom_bar(stat="identity", fill = "steelblue") + labs(x = 'Group', y = 'Average Score', caption = 'Red line indicates average score across all departments') + geom_hline(yintercept = total_average, col = "red", linetype="dashed")


ALLGroups %>% group_by(DESCRIP_E) %>% count()
```





